### 2019-10-18 manuscript/sci-rep/reviewer-letter/reviewer-response.tex
added in response to reviewers

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{changepage}% http://ctan.org/pkg/changepage
\usepackage{float}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{graphicx}
\usepackage{ragged2e}
\usepackage{scrextend}
\usepackage{lastpage}
\pagestyle{fancy}
\renewcommand{\headrulewidth}{0pt}
\rhead{\includegraphics[height= 1cm]{ento1.png}\vspace{1em}}
\chead{\includegraphics[width = 5cm]{agritech1.png}}
\lhead{\includegraphics[width=3cm]{acet1.png}}

\fancyfoot{}
\setlength\headheight{40.8pt}
\RequirePackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{titlesec}
\titleformat{\section}{\normalfont\fontsize{10}{15}\bfseries}{\thesection}{0em}{}
\titlespacing{\section}{}{1.24em}{0.24em}

\cfoot{Page \thepage \hspace{1pt} of \pageref{LastPage}}

\begin{document}

\begin{addmargin}[2.8in]{}
Willett, Filgueiras, \newline
Benda, Zhang, and Kenworthy \newline
----------------------------- \newline
Denis S. Willett\newline
Department of Entomology \newline
Cornell AgriTech  \newline
Cornell University \newline
15 Castle Creek Drive \newline
Geneva, NY 14456 \newline
deniswillett@cornell.edu \newline
\end{addmargin}
\setlength{\parindent}{0cm}

October 24, 2019

\vspace{1.24em}

Scientific Reports

\vspace{1.24em}

Dear Dr. Geary,

\vspace{0.48em}
\setlength{\parindent}{1.24cm}

Thank you for your consideration of \textit{“Sting nematodes modify metabolomic profiles of host plants”} for publication in Scientific Reports.  We have revised the manuscript in accordance with reviewer recommendations.  A complete and detailed account of the revisions is included below.  The reviewer comments were very helpful, much appreciated, and have resulted in a better manuscript.  In addition to submitting a revised version of the manuscript incorporating the latest changes, we are also submitting production quality images.

\vspace{2em}

Much appreciated, \newline

\vspace{1em}

Denis S Willett
\vspace{0.48em}

Camila Cramer Filgueiras
\vspace{0.48em}

Nicole D Benda
\vspace{0.48em}

Jing Zhang
\vspace{0.48em}

Kevin E Kenworthy

\hline

\paragraph{Reviewer 1 (Remarks to the Author)}

You spend a lot of time talking about relative tolerance of the bermudagrass lines to sting nematode. However, I was never very clear on how you are defining tolerance or how this was determined.

On page 2 under Bermudagrass response to sting nematode feeding...I think you should use the term nematode population densities instead of "nematode densities" or "differences in nematode population."

Very well written and very interesting paper

Reviewer #2 (Technical Comments to the Author):

This is a well researched study and the manuscript well structured and written. The methods used are scientifically sound, results appropriately interpreted and discussed. Except for a few minor comments below, I recommend publication of this manuscript.

Line 58. The authors stated that the canonical correspondence analysis explained 42\% of the observed inertia in the total metabolome. I suggest that the authors explain this because in Figure 2 this analysis explained 71.9\% of the total metabolome variation (CCA1: 45.3\% + CCA2: 26.6\%).

Lines 146 and 150. Change ‘conetainers’ to ‘containers and container’

I suggest that instead of putting in the appendix, the package names used for their statistical analysis, the authors should include these names in the text to educate readers, especially for those not familiar with the ‘R software’

Reviewer #2 (Remarks to the Author):

This is a well researched study and the manuscript well structured and written. The methods used are scientifically sound, results appropriately interpreted and discussed. Except for a few minor comments below, I recommend publication of this manuscript.

Line 58. The authors stated that the canonical correspondence analysis explained 42\% of the observed inertia in the total metabolome. I suggest that the authors explain this because in Figure 2 this analysis explained 71.9 \% of the total metabolome variation (CCA1: 45.3\% + CCA2: 26.6\%).

Lines 146 and 150. Change ‘conetainers’ to ‘containers and container’

I suggest that instead of putting in the appendix, the package names used for their statistical analysis, the authors should include these names in the text to educate readers, especially for those not familiar with the ‘R software

\end{document}